---
title: "P8105_hw3_qx2222"
author: "Qiaoyi Xu"
date: "2022-10-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1(answer posted)

##my answer:
```{r}
library(p8105.datasets)
data("instacart")

```

How many aisles are there, and which aisles are the most items ordered from?
There are `r nrow(distinct(instacart, aisle))` aisles. fresh fruits and fresh vegetables are the most items ordered from. 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle)%>%
  summarise(n_each_aisle = n()) %>%
  filter(n_each_aisle >= 10000) %>%
  arrange(desc(n_each_aisle)) %>%
  mutate(ordernumber_level = if_else(n_each_aisle <= 20000, "<=20k",
                                     if_else(n_each_aisle <= 30000 & n_each_aisle > 20000, "20k-30k",
                                             if_else(n_each_aisle <= 40000 & n_each_aisle > 30000, "30k-40k",
                                                     if_else(n_each_aisle <= 50000 & n_each_aisle > 40000, "40k-50k",
                                                             if_else(n_each_aisle > 50000, ">50k", NA_character_))))))%>%
  mutate(ordernumber_level = factor(ordernumber_level, levels = c(">50k", "40k-50k", "30k-40k", "20k-30k","<=20k")))%>%
  ggplot(aes(x = aisle, y = n_each_aisle)) +
  geom_col() +
  facet_wrap(vars(ordernumber_level), ncol = 1) +
  labs(title = "number of orders on each item", y = "number of orders") +
  theme_bw(base_size = 5) +
  theme(axis.text.x = element_text(face="bold", color="#993333", size=6, angle=45))

```

##answer:

#### Read in the data

```{r}
instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```





### Problem 2
1. Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. 


```{r}
accel_data = read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(cols = "activity_1":"activity_1440", names_to = "time", values_to = "activities" )%>%
  separate(time, c("repeatedword","minute"), sep = "_") %>%
  select(-repeatedword) %>%
  arrange(week, day) %>%
  mutate(minute = as.integer(minute)) %>%
  mutate(weekday_or_weekend = if_else(day %in% c("Saturday", "Sunday"), "weekend", "weekday" )) %>%
  mutate(weekday_or_weekend = factor(weekday_or_weekend))

accel_data

```

Describe the resulting dataset:

This cleaned dataset has `r nrow(accel_data)` observations and `r ncol(accel_data)` varibales, such as `r names(accel_data)`. In those varibales, week,day_id, day, minute shows the exact time of collected data. variable activities shows the number of activity happened in each time.  Finally, weekday_or_weekend variable we created shows data collected in weekday or weekend.


2. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
totals_table  = accel_data %>% group_by(day_id) %>%
  mutate(total = sum(activities))

totals_table
```

```{r}
totals_table %>%
  ggplot(aes(x = day_id, y = total, color = weekday_or_weekend )) + 
  geom_point() +
  labs(x = "Day_id", y = "total activities", title = "Total number of activities each day", 
       color = "weekday/weekend") +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.title = element_blank())

```


From the plot we got, we could see the line showed total activities counts on weekday is more stable, compared with the line of weekend. Besides, for the total activities counts on weekday has sightly increasing trend. But for the total activities counts on weekend has sightly decreasing trend.


3. Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r}
accel_data %>%
  mutate(hour = minute/60) %>%
  ggplot(aes(x = hour, y = activities, color = day)) +
  geom_point() +
  geom_line(alpha = .5) +
  geom_smooth(aes(group = day), se = FALSE) +
  labs(x = "Hour", y = "number of activities", title = "the 24-hour activity time courses for each day", 
       color = "Day of week")
```

From this single-panel plot that shows the 24-hour activity time courses for each day, we could see lower total activities counts on morning time (before 5am) each day and higher total activities counts on evening time (around 6pm-10pm). Then, from the smooth lines, it also shows total activities counts on evening time is greater than other time on most of days, especially on Friday. 


### Problem 3
```{r}
#load data of problem 3
library(p8105.datasets)
data("ny_noaa")
```


To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):



1. Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?


```{r}
clean_noaa = ny_noaa %>%
  separate(date, c("year","month","day"), sep = "-") %>%
  mutate(year = as.numeric(year),
         month = as.numeric(month),
         day = as.numeric(day),
         tmax = as.numeric(tmax)/10,
         tmin = as.numeric(tmin)/10,
         prcp = as.numeric(prcp)/10)

clean_noaa
```



```{r}
snowball_values = ny_noaa %>%
  group_by(snow) %>%
  summarise(snowball_obs = n())%>%
  arrange(desc(snowball_obs))

snowball_values
```


2. Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
average_maxtemp = clean_noaa %>%
  mutate(month = month.name[month]) %>%
  filter(month %in% c ("January", "July")) %>%
  group_by(id,year,month) %>%
  summarise(mean_tmax = mean(tmax, na.rm = T)) %>%
  filter(!is.na(mean_tmax)) %>%#clean missing data
  ggplot(aes(x = year, y = mean_tmax))+
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(.~month) +
  theme(panel.spacing = unit(1,"lines")) %>%
  labs(x = "Year", y = "Average Max Temperature",
       title = "the average max temperature in January and in July in each station")

average_maxtemp 

```


3. Make a two-panel plot showing 
(i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 

```{r}
tmax_vs_tmin = clean_noaa %>%
  ggplot(aes(x = tmin, y = tmax))+
  geom_hex() +
  labs(x = "Minimum temperature (tenths of degrees C)", y = "Maximum temperature (tenths of degrees C)", 
       title = "tmax vs tmin for the full dataset", fill = "counts") +
  scale_x_continuous(n.breaks = 6) +
  scale_fill_continuous(breaks = c(10000,30000,50000)) +
  theme_minimal() +
  theme(legend.position = "bottom")

tmax_vs_tmin
  
```



(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
dist_snowfall = clean_noaa %>%
  filter(snow>0&snow<100) %>%
  mutate(year = factor(year)) %>%
  ggplot(aes(x = snow, y = year, fill = year))+
  geom_density_ridges(alpha = .5, scale = 2) +
  labs(x = "Snowfall(mm)", y = "year", title = "the distribution of snowfall values by year")

dist_snowfall  
```

Finally, combine this two plot into a two-panel plot.


```{r}
final_two_panel = tmax_vs_tmin + dist_snowfall 

final_two_panel
```













